---
title: " Bulk RNA-Sequencing REMI Tutorial"
output: html_document
author: "Alice Yu"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The files that are loaded need to be downloaded from the Github
Reading in required libraries
Reading in code and databases

```{r setup, include=FALSE}
# Loading packages
library(remi)
```

Input file is a log-transformed bulk flow-sorted RNA-sequencing dataset
Data is from Gentles et al. 2020
https://genomebiology.biomedcentral.com/articles/10.1186/s13059-020-02019-x 
Column names are labeled as samplename_celltype

```{r}
# this should be a dataset that the user can download on their own for the tutorial?
dat <- read.csv("~/Projects/rotation_remi/data/test/bulk_flowsort/U01Adeno_logged.csv", row.names=1)
```

Set up data for REMI
There is a built in filtering step for REMI where users can choose to threshold their genes based on quantiles. The number 4 represents binning the data into four quantiles and removing data with expression levels in the lowest quantile. Filtering can also be turned off (filter = F)

OR users can pre-threshold the genes and turn the filter off, but setupData needs to run in order to make a REMI object for the algorithm

Data must be log-transformed for input, no prior scaling or normalization

```{r}
# list of cell types user wants to analyze 
celltypes <- c("I", "E", "F", "M")
cellexp.list <- setupData(dat, celltypes, 4, filter=T)

# RUNNING REMI
adeno.remi <- remi(celltypes, cellexp.list, cd="Louvain",
                   maxNum = ncol(cellexp.list$filtered$I))
```







